---
title: "Reservation Cancellation Episode7"
output: html_document
date: "2023-01-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages({
library(tidyverse) # metapackage of all tidyverse packages
library(tidymodels) # metapackage see https://www.tidymodels.org/
library(finetune)
library(timeDate)
library(bonsai)
library(discrim)
library(corrr)
library(GGally)
library(stacks)  # a custom ensembling tool
})
    
tidymodels_prefer()
conflicted::conflicts_prefer(recipes::update)
conflicted::conflicts_prefer(purrr::is_null)

options(tidymodels.dark = TRUE)

theme_set(theme_bw())

metrics <- metric_set(roc_auc)

```


```{r}
if (dir.exists("/kaggle")){
  path <- "/kaggle/input/playground-series-s3e7/"
} else {
  path <- str_c(here::here("data"),"/")
}
```


```{r }
#| label: load data

factors <- c("room_type_reserved", "required_car_parking_space", "type_of_meal_plan", "market_segment_type", "repeated_guest","holiday")

preprocessor <- function(data){
  data |> 
    janitor::clean_names() |> 
    mutate(idxRelevance = ((no_of_special_requests + required_car_parking_space)
                    *(no_of_weekend_nights + no_of_week_nights)
                    *avg_price_per_room*(no_of_adults + no_of_children))*(repeated_guest + 1)) |> 
    mutate(arrival_date = 
      case_when(arrival_month == 2 & arrival_date %in% c(29,30,31) ~ 28,
                arrival_month %in% c(2,4,6,9, 11) & arrival_date == 31 ~ 30,
                .default = arrival_date
                  ),
    arrival = lubridate::ymd(paste0(arrival_year,"-",arrival_month,"-",arrival_date)),
    holiday = if_else(timeDate::isHoliday(timeDate(arrival)), "holiday","no_holiday"),
    booking = arrival - lubridate::days(lead_time),
    across(all_of(factors), factor)
    ) |> 
    select(- arrival_date, 
           - arrival_month, 
           - arrival_year)
}

raw_df <- read_csv(str_c(path, "train.csv"),
                   show_col_types = FALSE) |> 
          preprocessor() |> 
          distinct(across(-id), .keep_all = TRUE) |> 
          mutate(booking_status = factor(booking_status,
                                         labels = c("stayed","cancelled")))

competition_df <- read_csv(str_c(path, "test.csv"),
                   show_col_types = FALSE) |> 
          preprocessor()

all_df <- bind_rows(
  raw_df |> mutate(source = "train"),
  competition_df |> mutate(source = "competition")
) |> mutate(source = factor(source))

```

Booking_ID: unique identifier of each booking

No of adults: Number of adults
No of children: Number of Children
noofweekend_nights: Number of weekend nights (Saturday or Sunday) the guest stayed or booked to stay at the hotel
noofweek_nights: Number of week nights (Monday to Friday) the guest stayed or booked to stay at the hotel
typeofmeal_plan: Type of meal plan booked by the customer
requiredcarparking_space: Does the customer require a car parking space? (0 - No, 1- Yes)
roomtypereserved: Type of room reserved by the customer. The values are ciphered (encoded) by INN Hotels.
lead_time: Number of days between the date of booking and the arrival date
arrival_year: Year of arrival date
arrival_month: Month of arrival date
arrival_date: Date of the month
Market segment type: Market segment designation.
repeated_guest: Is the customer a repeated guest? (0 - No, 1- Yes)
noofprevious_cancellations: Number of previous bookings that were canceled by the customer prior to the current booking
noofpreviousbookingsnot_canceled: Number of previous bookings not canceled by the customer prior to the current booking
avgpriceper_room: Average price per day of the reservation; prices of the rooms are dynamic. (in euros)
noofspecial_requests: Total number of special requests made by the customer (e.g. high floor, view from the room, etc)
booking_status: Flag indicating if the booking was canceled or not.


Is this competition transaction already in the training data with a correct label?

```{r}

all_df |> 
  group_by(-id) |>  
  mutate(num_dups = n(), 
         dup_id = row_number()) |>  
  ungroup() |> 
  group_by(source, booking_status) |> 
  mutate(is_duplicated = dup_id > 1) |> 
  count(is_duplicated)

```

This dataset appears to lack any duplicates, after applying the pre-processor.

Lets zoom in more closely on the training set only.


```{r}
#| label: numeric density plots

raw_df |> 
  select(where(is.numeric), booking_status) |> 
  pivot_longer(- c(id, booking_status),
    names_to = "metric",
    values_to = "value"
  ) |> 
  filter(value > 0.1) |> 
  ggplot(aes(value, fill = factor(booking_status), color = NULL)) +
  geom_density(alpha = 0.6, show.legend = FALSE) +
  scale_fill_brewer(palette = "RdYlBu") +
  facet_wrap(vars(metric, booking_status), scales = "free", ncol = 3) +
  theme_dark() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())


all_df |> 
  select(where(is.numeric), booking_status) |> 
  pivot_longer(- c(id, booking_status),
    names_to = "metric",
    values_to = "value"
  ) |> 
  ggplot(aes(sample = value, color = booking_status)) + 
  stat_qq(show.legend = FALSE) + 
  stat_qq_line(show.legend = FALSE) +
  scale_color_brewer(palette = "RdYlBu") +
  facet_wrap(vars(metric), scales = "free", ncol = 3) +
  theme_dark() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())




raw_df %>% 
  drop_na() %>% 
  select(all_of(factors), booking_status) %>% 
  pivot_longer(c(-booking_status), names_to = "type", values_to = "value") %>% 
  ggplot(aes(value, fill = booking_status)) +
  geom_bar(position = "dodge") +
  facet_wrap(~ type, scales = "free", nrow = 2) +
  theme_minimal() +
  theme(legend.position = "top") +
  labs(title = "Target impact - categorical features - step 1")

raw_df %>% 
  drop_na() %>% 
  select(all_of(factors), booking_status) %>% 
  pivot_longer(c(-booking_status), names_to = "type", values_to = "value") %>% 
  group_by(type, value, booking_status) %>% 
  summarize(n = n(),
            .groups = "drop") |> 
  mutate(frac = n / sum(n)) |> 
  filter(booking_status != FALSE) %>% 
  ggplot(aes(value, frac, fill = type)) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~ type, scales = "free_x") +
  theme_minimal() +
  theme(legend.position = "top") +
  labs(title = "Target impact - categorical features - step 2: percentages of hotel visits")

# a function to extract binomial confidence levels, by kaggle user headsortails
get_binCI <- function(x,n){
  test_result <- binom.test(x,n)
  test_result_ci <- test_result$conf.int
  test_result_ci <- setNames(test_result_ci, c("lwr", "upr"))
  return(as.list(test_result_ci))
}

raw_df %>% 
  drop_na() %>% 
  select(all_of(factors), booking_status) %>% 
  pivot_longer(c(-booking_status), names_to = "type", values_to = "value") %>% 
  count(type, value, booking_status) %>% 
  pivot_wider(names_from = booking_status, values_from = n) %>%
  group_by(type, value) %>% 
  mutate(frac = stayed/(stayed + cancelled)*100,
         lwr = get_binCI(stayed,(stayed + cancelled))[[1]]*100,
         upr = get_binCI(stayed,(stayed + cancelled))[[2]]*100
         ) %>%
  ggplot(aes(value, frac, col = type)) +
  geom_point(size = 4, show.legend = FALSE) +
  geom_errorbar(aes(ymin = lwr, ymax = upr), show.legend = FALSE) +
  facet_wrap(~ type, scales = "free_x") +
  labs(title = "Target impact - categorical features - step 3: points with uncertainties")


```



```{r }
#| label: feature target interaction
#| fig.height: 4.5
raw_df %>% 
  drop_na() %>% 
  group_by(room_type_reserved, required_car_parking_space, booking_status) %>% 
  summarise(avg_price_per_room = median(avg_price_per_room, na.rm = TRUE),
            .groups = "drop") %>% 
  ggplot(aes(room_type_reserved, required_car_parking_space, fill = avg_price_per_room)) +
  geom_tile() +
  scale_fill_viridis_c() +
  facet_wrap(~ booking_status, nrow = 2) +
  theme_minimal() +
  theme(legend.position = "right")
```


```{r }
#| label: feature target interaction
#| fig.height: 4.5
raw_df %>% 
  drop_na() %>% 
  group_by(market_segment_type, repeated_guest, booking_status) %>% 
  summarise(avg_price_per_room = median(avg_price_per_room, na.rm = TRUE),
            .groups = "drop") %>% 
  ggplot(aes(market_segment_type, repeated_guest, fill = avg_price_per_room)) +
  geom_tile() +
  scale_fill_viridis_c() +
  facet_wrap(~ booking_status, nrow = 2) +
  theme_minimal() +
  theme(legend.position = "right")
```


```{r }
#| label: feature target interaction4
#| fig.height: 4.5

plot_interaction <- function(data, var1, var2){

data %>% 
  drop_na() %>% 
  group_by({{var1}} := ggplot2::cut_number({{var1}}, 5), {{var2}} := ggplot2::cut_number({{var2}}, 5), booking_status) %>% 
  summarise(avg_price_per_room = median(avg_price_per_room, na.rm = TRUE),
            .groups = "drop") %>% 
  ggplot(aes({{var1}}, {{var2}}, fill = avg_price_per_room)) +
  geom_tile() +
  scale_fill_viridis_c() +
  facet_wrap(~ booking_status, nrow = 2) +
  theme_minimal() +
  theme(legend.position = "right")
  
}

raw_df |> 
  transmute(arrival_week = lubridate::week(arrival),
         booking_week = lubridate::week(booking),
         avg_price_per_room, booking_status) %>%  
   plot_interaction(arrival_week, booking_week)

```

```{r }
#| label: feature target interaction5
#| fig.height: 4.5

raw_df |> 
  transmute(lead_time,
         booking_week = lubridate::week(booking),
         avg_price_per_room, booking_status) %>%  
   plot_interaction(lead_time, booking_week)



```


```{r}
#| label: make resample folds

set.seed(42)

folds <- vfold_cv(raw_df, 
                  v = 11,
                  repeats = 2,
                  strata = booking_status)

```

Too speed computation, we will load up a parallel backend.

```{r}
#| label: assign a parallel backend

all_cores <- parallelly::availableCores(omit = 1)
all_cores

future::plan("multisession", workers = all_cores) 
```


```{r}
#| label: model specifications

boost_tree_xgboost_spec <-
  boost_tree(
    trees = 2800L,
    min_n = 4L,  # 13L, # 7L, # 7, # 7L, 
    mtry = 22L,  # 23L, # 11L, 
    tree_depth = 4L, # 6L, # 8L# 10L,  
    learn_rate = 0.0192,# 0.1, 
    loss_reduction = 0, # 0,
    sample_size = 1, # ,
    stop_iter = 11L # 9L #16L
  ) |> 
  set_engine('xgboost', 
          #   validation = 0.1,
             eval_metric = "auc",
             objective = "binary:logistic",
             event_level = "second") |> 
  set_mode('classification')

boost_tree_lgbm_spec <- 
  boost_tree(
    trees = 800L,
    tree_depth = 20L,
    learn_rate = 0.03,
    mtry = 26L,  
    min_n = 19L
    # loss_reduction = 0
  ) %>% 
  set_engine(engine = "lightgbm") %>%
  set_mode(mode = "classification") 

rand_forest_ranger_spec <-
  rand_forest(
    mtry = 8L,# 13L, 
    min_n = 40L, #46L
              ) %>%
  set_engine('ranger', importance = "impurity") %>%
  set_mode('classification')

rand_forest_randomForest_spec <-
  rand_forest(
    mtry = 11L, #13L
    min_n = 39L) %>%
  set_engine('randomForest') %>%
  set_mode('classification')

multinom_reg_glmnet_spec <-
  logistic_reg(
    penalty = 1.78e-4,# 1e-3, 
    mixture = 1.0
    ) %>%
  set_engine('glmnet')

discrim_flexible_earth_spec <-
  discrim_flexible(
    num_terms = 41L, #37,
    prod_degree = 2
  ) %>%
  set_engine('earth')

rec <- recipe(booking_status ~ .,
         data = raw_df ) |>
  update_role(id, new_role = "id") |>
  step_date(arrival, booking,
            ordinal = TRUE,
            features = c("dow", "week","month"),
            keep_original_cols = FALSE) |> 
  step_mutate(total_people = no_of_adults + no_of_children) |> 
  step_mutate(total_nights = no_of_weekend_nights + no_of_week_nights) |> 
  step_mutate(total_previous_bookings = no_of_previous_cancellations + no_of_previous_bookings_not_canceled) |> 
  step_mutate(prc_cancellations = if_else(total_previous_bookings == 0, 0, no_of_previous_cancellations/total_previous_bookings )) |> 
  step_mutate(price_per_person = if_else(total_people == 0, 0, avg_price_per_room / total_people)) |> 
  step_mutate(total_price = avg_price_per_room * total_nights) |> 
  step_interact(~ arrival_week:booking_week) |> 
  step_interact(~ lead_time:booking_week) |> 
  step_interact(~ lead_time:total_price) |> 
  step_bs(lead_time) 


```


```{r}
#| label: any other interesting feature interactions
#| warning: false
#| message: false
#| fig.height: 10
#| fig.width: 12

foo <- rec |> prep() |> bake(new_data = NULL) %>% 
  select(where(is.numeric), -id, booking_status) %>% 
  drop_na() 

foo %>% 
  ggpairs(
    columns = 1:(ncol(foo)-1),
    mapping = aes(color = booking_status, alpha = 0.5),
    lower = list(continuous = wrap("points", alpha = 0.3, size=0.01)),
    upper = list(continuous = wrap("smooth", alpha = 0.005, size = 0.1)),
    progress = FALSE) +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  labs(title = "Pair plots: lower: scatter; upper: linear fit - color by target")
```


```{r}
# xgboost_param <- boost_tree_xgboost_spec |>
#   extract_parameter_set_dials() |>
#   update(
#     min_n = min_n(range = c(3L, 16L)),
#     mtry = mtry(range = c(15L, 30L)),
#     tree_depth = tree_depth(range = c(6L, 12L))
#     # stop_iter = stop_iter(range = c(10,18))
#   ) |>
#   finalize(rec |> prep() |> bake(new_data = NULL))
# 
# xgb_grid <-
#    grid_latin_hypercube(xgboost_param, size = 45)

# lgbm_param <- boost_tree_lgbm_spec |>
#   extract_parameter_set_dials() |>
#   update(
#     tree_depth = tree_depth(range = c(18, 23)),
#     min_n = min_n(range = c(15L, 21L)),
#      mtry = mtry(range = c(20, 28))
#     # tree_depth = tree_depth(range = c(10L, 14L)),
#   ) |>
#   finalize(rec |> prep() |> bake(new_data = NULL))

# 
# rand_forest_ranger_param <- rand_forest_ranger_spec %>%
#   extract_parameter_set_dials() |>
#   update(
#     mtry = mtry(),
#     min_n = min_n(range = c(6L, 50L))
#   ) |>
#   finalize(rec |> prep() |> bake(new_data = NULL))
# 
# rf_grid <- 
#    grid_regular(rand_forest_ranger_param, levels = c(9, 5))
# 

# 
# 
# multinom_reg_glmnet_param <- multinom_reg_glmnet_spec %>%
#   extract_parameter_set_dials() %>%
#   update(
#     penalty = penalty(range = c(-4.5,-3.5)), 
#     mixture = mixture()) 
# 
# glmn_grid <- 
#   grid_regular(multinom_reg_glmnet_param, levels = c(9, 5))

```



```{r}
#| label: xgb
ctrl <- control_resamples(
     verbose = FALSE,
     save_pred = TRUE,
     save_workflow = TRUE,
     parallel_over = "everything")

ctrlg <- finetune::control_race(
     verbose = TRUE,
     verbose_elim = TRUE,
     burn_in = 2,
     save_pred = TRUE,
     save_workflow = TRUE,
     parallel_over = "everything")

# results <- fit_resamples(
#   workflow(rec |> step_dummy(all_nominal_predictors(),
#                              one_hot = TRUE), 
#            boost_tree_xgboost_spec),
#   resamples = folds,
#   #  grid = 15,
#   control = ctrl,
#   metrics = metrics,
#   #  param_info = xgboost_param
# )


boost_tree_xgboost_spec <-
  boost_tree(
    trees = 2800L,
    min_n = tune(), # 10L,  # 13L, # 7L, # 7, # 7L,
    mtry =tune(), #  16L,  # 23L, # 11L,
    tree_depth = tune(), #4L, # 6L, # 8L# 10L,
    learn_rate = tune(), # 0.1,# 0.1,
    loss_reduction = 0, # 0,
    sample_size = 1, # ,
    stop_iter = tune(), # 5L # 9L #16L
  ) |>
  set_engine('xgboost',
          #   validation = 0.1,
             eval_metric = "auc",
             objective = "binary:logistic",
             event_level = "second") |>
  set_mode('classification')

xgboost_param <- boost_tree_xgboost_spec |>
  extract_parameter_set_dials() |>
  update(
    min_n = min_n(range = c(3L, 8L)),
    mtry = mtry(range = c(20L, 25L)),
    tree_depth = tree_depth(range = c(3L, 6L)),
    learn_rate = learn_rate(),
    stop_iter = stop_iter(range = c(5,16))
  ) |>
  finalize(rec |> prep() |> bake(new_data = NULL))

results <- finetune::tune_race_anova(
  workflow(rec |>
             step_dummy(all_nominal_predictors())|>
             step_range(all_numeric_predictors()),
                        boost_tree_xgboost_spec),
  resamples = folds,
  grid = 10,
  control = ctrlg,
  metrics = metrics,
  param_info = xgboost_param
  )

autoplot(results)

results |> 
  collect_metrics(summarize = TRUE) |> 
  arrange(desc(mean))

workflow(rec |> step_dummy(all_nominal_predictors()),
                        boost_tree_xgboost_spec) %>%
#  finalize_workflow(select_best(results)) %>% 
   fit(raw_df) %>%
   extract_fit_parsnip() %>%
   vip::vip(num_features = 20) +
  labs(title = "XGBoost Variable Importance")
```



```{r}
#| label: lgbm
results2 <- fit_resamples(
  workflow(rec, boost_tree_lgbm_spec),
  resamples = folds,
  control = ctrl,
  metrics = metrics
)



results2 |> 
  collect_metrics(summarize = TRUE) |> 
  arrange(desc(mean))

workflow(rec, boost_tree_lgbm_spec) %>%
fit(raw_df) %>%
  extract_fit_engine() %>%
   lgb.importance(percentage = TRUE) %>%
   lgb.plot.importance(top_n = 20) 
```

```{r}
#| label: ranger
results3 <- fit_resamples(
  workflow(rec, rand_forest_ranger_spec),
  resamples = folds,
  control = ctrl,
  metrics = metrics
)


results3 %>% 
  collect_metrics(summarize = TRUE) %>%
  arrange(desc(mean))

workflow(rec, rand_forest_ranger_spec) %>%
   fit(raw_df) %>%
   extract_fit_parsnip() %>%
   vip::vip(num_features = 20) +
  labs(title = "Ranger Variable Importance")
```

```{r}
#| label: randomForest
results4 <- fit_resamples(
  workflow(rec, rand_forest_randomForest_spec),
  resamples = folds,
  control = ctrl,
  metrics = metrics
)


results4 %>% 
  collect_metrics(summarize = TRUE) %>%
  arrange(desc(mean))

workflow(rec, rand_forest_randomForest_spec) %>%
   fit(raw_df) %>%
   extract_fit_parsnip() %>%
   vip::vip(num_features = 20) +
  labs(title = "Random Forest Variable Importance")

```


```{r}
#| label: glmnet
results5 <- fit_resamples(
  workflow(rec |> step_dummy(all_nominal_predictors()) |> 
  step_normalize(all_numeric_predictors()) ,
  multinom_reg_glmnet_spec),
  resamples = folds,
  control = ctrl,
  metrics = metrics
)

results5 %>% 
  collect_metrics(summarize = TRUE) %>%
  arrange(desc(mean))

workflow(rec |> 
           step_dummy(all_nominal_predictors()) |> 
           step_normalize(all_numeric_predictors()),
         multinom_reg_glmnet_spec) %>%
   fit(raw_df) %>%
   extract_fit_parsnip() %>%
   vip::vip(num_features = 20) +
  labs(title = "GLMNet Variable Importance")

```



```{r}
#| label: descrim

results6 <- fit_resamples(
  workflow(rec |>step_dummy(all_nominal_predictors()) |> 
             step_corr(all_numeric_predictors()) |> 
             step_normalize(all_numeric_predictors()) , 
           discrim_flexible_earth_spec),
  resamples = folds,
  control = ctrl,
  metrics = metrics
)

results6 %>% 
  collect_metrics(summarize = TRUE) %>%
  arrange(desc(mean))

workflow(rec |> step_dummy(all_nominal_predictors()) |>  
           step_corr(all_numeric_predictors()) |> 
           step_normalize(all_numeric_predictors()) , 
         discrim_flexible_earth_spec) |> 
  fit(raw_df) |> 
  extract_fit_engine() |> 
  mda::coef.fda()  


```


```{r}
#| label: stacks
ens <- stacks() |>
  stacks::add_candidates(results) |>
  stacks::add_candidates(results2) |>
  stacks::add_candidates(results3) |>
  stacks::add_candidates(results4) %>% 
  stacks::add_candidates(results5) %>%
  stacks::add_candidates(results6) |> 
  stacks::blend_predictions(
    metric = metric_set(roc_auc),
    penalty = c(seq(0.001, 0.05, 0.001)),
    mixture = c(seq(0.6, 0.95, 0.05,)),
    non_negative = FALSE,
    control = tune::control_grid(allow_par = TRUE,
                                 event_level = "second")
  ) 

autoplot(ens)

autoplot(ens, "weights")

ensemble <- fit_members(ens)

```


```{r}

predict(ensemble, raw_df, type = "class") |> 
  bind_cols(raw_df) |> 
  conf_mat(booking_status, .pred_class) |> 
  autoplot("heatmap") +
  labs(title = "Training Data Ensemble")

predict(ensemble, raw_df, type = "prob") |> 
  bind_cols(raw_df) |> 
  roc_auc(booking_status, .pred_stayed)

```


```{r}
submission_df <- predict(ensemble, 
        competition_df, 
        type = "prob") %>%
  bind_cols(competition_df) %>% 
  select(id, booking_status = .pred_cancelled)

submission_df |> 
  ggplot(aes(booking_status)) +
  geom_histogram(bins = 40)

submission_df %>%
  write_csv(str_c(path, "submission.csv"))
```




