---
title: "Predict hits with NEW MLB swing data"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
    highlight: tango
    code_folding: hide
---
  
# Introduction  {.tabset .tabset-fade .tabset-pills}

The goal of this competition is to predict hits with NEW MLB swing data.

My notebook serves as a demonstration of some of the possible techniques available to arrive at a solution.  I intend to add to this as I have time available. Your questions and comments are welcome.

If you fork this on kaggle, be sure to choose the kernel Environment setting for "Always use latest environment"

Lets dive right in.

The Kaggle kernels have many of the common r packages built in.  

## Load libraries

In addition to `tidymodels` we will load the `bonsai` interface to lightgbm.

```{r }
#| label: setup
#| warning: false
#| message: false

if (dir.exists("/kaggle")){
  path <- "/kaggle/input/nwds-batted-balls/"
    
options(repos = c(CRAN = "https://packagemanager.posit.co/cran/2021-03-22"))
# install.packages("vip", quiet = TRUE)
    
cores <- future::availableCores()

} else {
  path <- stringr::str_c(here::here("data"),"/")

cores <- future::availableCores(omit = 1)
}

 
suppressPackageStartupMessages({
library(tidyverse, quietly = TRUE) # metapackage of all tidyverse packages
library(tidymodels) # metapackage see https://www.tidymodels.org/
library(bonsai)
})

tidymodels_prefer()

options(tidymodels.dark = TRUE)

theme_set(cowplot::theme_minimal_grid())

```


## Load Data

```{r }
#| label: load data
#| warning: false
#| message: false


train_spec <- cols(
  uid = col_double(),
  pitch_type = col_character(),
  release_speed = col_double(),
  release_pos_x = col_double(),
  release_pos_z = col_double(),
  game_type = col_character(),
  is_lhp = readr::col_factor(),
  is_lhb = readr::col_factor(),
  balls = readr::col_factor(),
  strikes = readr::col_factor(),
  game_year = readr::col_factor(),
  pfx_x = col_double(),
  pfx_z = col_double(),
  plate_x = col_double(),
  plate_z = col_double(),
  on_3b = readr::col_factor(),
  on_2b = readr::col_factor(),
  on_1b = readr::col_factor(),
  outs_when_up = readr::col_factor(),
  inning = readr::col_factor(),
  is_top = readr::col_factor(),
  vx0 = col_double(),
  vy0 = col_double(),
  vz0 = col_double(),
  ax = col_double(),
  ay = col_double(),
  az = col_double(),
  sz_top = col_double(),
  sz_bot = col_double(),
  effective_speed = col_double(),
  release_spin_rate = col_double(),
  release_extension = col_double(),
  release_pos_y = col_double(),
  pitch_number = col_double(),
  pitch_name = readr::col_factor(),
  spin_axis = col_double(),
  spray_angle = col_double(),
  bat_speed = col_double(),
  swing_length = col_double(),
  outcome = col_character(),
  outcome_code = col_character()
)


competition_spec <- cols(
  uid = col_double(),
  pitch_type = col_character(),
  release_speed = col_double(),
  release_pos_x = col_double(),
  release_pos_z = col_double(),
  game_type = col_character(),
  is_lhp = readr::col_factor(),
  is_lhb = readr::col_factor(),
  balls = readr::col_factor(),
  strikes = readr::col_factor(),
  game_year = readr::col_factor(),
  pfx_x = col_double(),
  pfx_z = col_double(),
  plate_x = col_double(),
  plate_z = col_double(),
  on_3b = readr::col_factor(),
  on_2b = readr::col_factor(),
  on_1b = readr::col_factor(),
  outs_when_up = readr::col_factor(),
  inning = readr::col_factor(),
  is_top = readr::col_factor(),
  vx0 = col_double(),
  vy0 = col_double(),
  vz0 = col_double(),
  ax = col_double(),
  ay = col_double(),
  az = col_double(),
  sz_top = col_double(),
  sz_bot = col_double(),
  effective_speed = col_double(),
  release_spin_rate = col_double(),
  release_extension = col_double(),
  release_pos_y = col_double(),
  pitch_number = col_double(),
  pitch_name = readr::col_factor(),
  spin_axis = col_double(),
  spray_angle = col_double(),
  bat_speed = col_double(),
  swing_length = col_double()
)

preprocessor <- function(dataframe) {

dataframe <- dataframe %>%
    janitor::clean_names() %>%
  
    mutate(pitch_speed_diff = release_speed - effective_speed, # from Michael Semenoff notebook
           pitch_location = sqrt(plate_x^2 + plate_z^2),
           count_pressure = balls - strikes + 1,
           bat_pitch_speed_ratio = bat_speed / release_speed,
           swing_efficiency = bat_speed / swing_length) %>%
  
    mutate(across(c(where(is.character)), ~ as.factor(.x))) 

return(dataframe)
}

raw_df <- read_csv(str_c(path, "train.csv"),
                   col_types = train_spec,
                   show_col_types = FALSE) %>%
          preprocessor() 

tst_df <- read_csv(str_c(path, "test.csv"),
                   col_types = competition_spec,
                   show_col_types = FALSE)  %>% 
  preprocessor() 

all_df <-
    bind_rows(raw_df %>% mutate(source = "train"),
              tst_df %>% mutate(source = "test"))

train_df <- all_df %>% 
  filter(source == "train") %>% 
  select(-source, -outcome_code, -game_year) %>% 
  mutate(outcome = factor(outcome, levels = c("out","single","double","triple","home_run")))

competition_df <- all_df %>% 
  filter(source == "test") %>% 
  select(-source, -outcome, -outcome_code, -game_year)

features <- train_df %>%
  select(-uid, -outcome) %>%
  names()

train_df <- train_df %>% 
  distinct(pick(all_of(features)), .keep_all = TRUE)

nom_features <- train_df %>%
  select(all_of(features)) %>%
  select(where(is.character), where(is.factor)) %>%
  names() 

logical_features <- train_df %>%
  select(all_of(features)) %>%
  select(where(is.logical)) %>%
  names() 

num_features <- train_df %>%
  select(all_of(features)) %>%
  select(where(is.numeric)) %>%
  names()

bundle_split <- initial_split(train_df, prop = 0.95, strata = outcome)
bundle_train <- training(bundle_split)
bundle_test <- testing(bundle_split)


```

Nominal features:

`r nom_features`

Numeric features: 

`r num_features`

Logical features: 

`r logical_features`


Size of the combined train and competition datasets:

`r nrow(all_df)`

Size of the split made available to machine learning

`r nrow(bundle_train)`


# EDA {.tabset .tabset-fade .tabset-pills}

## Numeric features

Consider where features require univariate transformation, or clipping outliers.
```{r}
#| label: numeric
#| warning: false
#| message: false
#| fig.height: 24
#| fig.width: 6

train_df %>% 
  select(all_of(num_features), outcome ) %>% 
  pivot_longer(-outcome,
    names_to = "metric",
    values_to = "value"
  ) %>%
  ggplot(aes(value, fill = outcome)) +
  geom_histogram(bins = 200) +
   facet_wrap(vars(metric), scales = "free", ncol = 2) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "top", 
        plot.title.position = "plot") +
  labs(color = NULL, fill = NULL,
       title = "Numeric Feature Univariate Distributions",
       caption = "Data: MLB Statcast via Baseball Savant | Visual: Jim Gruman")

```


## Nominal features

Explore the distribution of outcome class by factor level, and the factor levels that exist in test that do not exist in training.


```{r}
#| label: nominal
#| warning: false
#| message: false
#| fig.height: 18
#| fig.width: 6


if(length(nom_features) >0){

all_df %>% 
  select(all_of(nom_features), outcome) %>% 
  mutate(across(nom_features, fct_lump_n,n = 10, other_level = 'other')) %>%
  pivot_longer(-outcome,
    names_to = "metric",
    values_to = "value"
  ) %>%
  
  summarise(n = n(),
            .by = c(outcome, metric, value)) %>%
      
  mutate(value = tidytext::reorder_within(value, n, metric)) %>%
    
  ggplot(aes(x = n, y = value, fill = outcome)) +
  geom_col() +
  tidytext::scale_y_reordered() +
  scale_x_continuous(n.breaks = 3, guide = guide_axis(n.dodge = 2))  +
  facet_wrap(vars(metric), scales = "free", ncol = 2) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
       legend.position = "bottom") +
  labs(title = "Nominal Feature Counts",
       fill = NULL,
       caption = "Data: MLB Statcast via Baseball Savant | Visual: Jim Gruman")

}

```

## Logical features

```{r}
#| label: logical
#| warning: false
#| message: false
#| fig.height: 6
#| fig.width: 6


if(length(logical_features) >0){

train_df %>% 
  select(all_of(logical_features), class) %>% 
  pivot_longer(-class,
    names_to = "metric",
    values_to = "value"
  ) %>%
  ggplot(aes(y = value, fill = class)) +
  geom_bar() +
  scale_x_continuous(n.breaks = 3, guide = guide_axis(n.dodge = 2)) +
   facet_wrap(vars(metric), scales = "free", ncol = 2) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
       legend.position = "bottom") +
  labs(title = "Logical Feature Counts",
       fill = "Class",
       caption = "Data: Kaggle.com | Visual: Jim Gruman")
}

```

## Counts of Missingness
                  
```{r}
#| label: counts of missingness

#naniar::gg_miss_upset(train_df,  nintersects = 6) 

#naniar::vis_miss(train_df |> slice_sample(prop = 0.01))

train_df %>% 
  summarize(across(all_of(features), function(x) sum(is.na(x)))) %>% 
  pivot_longer(everything(),
              names_to = "feature",
              values_to = "Count of Missing") %>% 
                   knitr::kable()
                   

                  
```

## Counts of Distinct
                   
               
```{r}
#| label: counts of distinct
               
raw_df %>%
  summarize(across(all_of(features), n_distinct)) %>%
  pivot_longer(everything(), names_to = "feature", values_to = "Count of distinct train") |>
  left_join(
    tst_df %>%
      summarize(across(all_of(features), n_distinct)) %>%
      pivot_longer(everything(), names_to = "feature", values_to = "Count of distinct test")
  ) %>% 
                   knitr::kable()
               
```

## Duplicated

Is this competition transaction already in the training data with a correct label?

```{r}
#| label: duplicates
#| warning: false
#| message: false
all_df %>%
    select(all_of(features), source) %>% 
    group_by_at(features) %>%
    mutate(num_dups = n(),
           dup_id = row_number()) %>%
    ungroup() %>%
    group_by(source) %>%
    mutate(is_duplicated = dup_id > 1) %>%
    count(is_duplicated) %>% 
                   knitr::kable()
               

```
                   



## Target

```{r}
#| label: outcome 
#| warning: false
#| message: false
#| fig.width: 6


bundle_train %>% 
  summarize(outcome_sum = n(),
            .by = outcome) %>%
  arrange(outcome_sum) %>%
  mutate(prop = outcome_sum / nrow(bundle_train)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop) %>%

  ggplot(aes(x = "", y = prop, fill = outcome)) +
  geom_bar(stat = "identity", width = 1, show.legend = FALSE) +
  geom_text(aes(y = ypos, label = paste0(outcome
                                  ,"\n",round(prop,2)*100,"%")),
            color = "white",
            nudge_x = 0.3,
            size = 3) +
  coord_polar("y", start = 0) +
  theme_void() + 
  labs(title = "Outcome",
       caption = "Data: Kaggle.com | Visual: Jim Gruman")

```
                              
           
                
# Machine Learning {.tabset .tabset-fade .tabset-pills}


## Recipe

```{r}
#| label: recipe
                   
rec <- recipe(
    
    formula(paste0("outcome ~ ", 
               str_c(features,  collapse = " + "))),
    data = bundle_train
  ) %>% 
  step_novel(all_nominal_predictors()) %>%
  step_nzv(all_predictors())
                   
folds <- vfold_cv(bundle_train, 
                  v = 3,
                  repeats = 1,
                  strata = outcome)    
                   
ctrl <- finetune::control_sim_anneal(     
     verbose = FALSE,
     verbose_iter = TRUE,
     parallel_over = "everything",
     save_pred = FALSE,
     save_workflow = FALSE)                   
                                     
```



## Lightgbm

```{r}
#| label: lgbm


boost_tree_lgbm_spec <- 
  boost_tree(
    trees = 1350L,
   tree_depth = tune(),
   learn_rate =  tune(),
   min_n = tune(),
   mtry = tune(),
   loss_reduction = 0
  ) %>% 
  set_engine(engine = "lightgbm",
             is_unbalance = TRUE,
             num_leaves = tune(),
             num_threads = cores
             ) %>%
  set_mode(mode = "classification") 
                   
wf <- workflow(rec,
               boost_tree_lgbm_spec) 

param <- wf %>%
   extract_parameter_set_dials() %>%
   recipes::update(
      min_n = min_n(range = c(5,20)),
      tree_depth = tree_depth(range = c(4,30)),
      learn_rate = learn_rate(range = c(-1.0,-1.6)),
      num_leaves = num_leaves(range = c(100,400))
   ) %>%
   dials::finalize(raw_df)                 

burnin <- tune_grid(
  wf,
  grid = 4,
  resamples = folds,
  control = ctrl,
  metrics = metric_set(roc_auc, accuracy),
  param_info = param)

lgbm_rs <- finetune::tune_sim_anneal(
  wf,
  resamples = folds,
  iter = 3,
  initial = burnin,
  control = ctrl,
  metrics =  metric_set(roc_auc, accuracy),
  param_info = param) 

show_best(lgbm_rs, metric = "roc_auc")  

autoplot(lgbm_rs)

collect_metrics(lgbm_rs) %>% 
  filter(.metric == "roc_auc") %>% 
  mutate(.config = fct_reorder(.config, mean)) %>% 
  ggplot(aes(mean, .config)) +
  geom_point() +
  geom_errorbarh(aes(xmin = mean - std_err, xmax = mean + std_err)) +
  labs(title = "ROC AUC across resample folds with std_err")
                   

```
                   
# Submission
                   
```{r}
#| label: submission
#| warning: false
#| message: false

highest_roc <-   select_best(lgbm_rs, metric = "roc_auc")                 
                   
final_lgbm <- wf %>%
    finalize_workflow(highest_roc) 

last_fit(
  final_lgbm,
  bundle_split,
  metrics = metric_set(roc_auc, accuracy)
) %>%
  collect_metrics() %>%
  knitr::kable()

classification_fit <- final_lgbm %>%
                        fit(bundle_train) 

classification_fit |> 
  extract_fit_engine() |> 
  lgb.importance() |> 
  as_tibble() %>% 
  ggplot(aes(x = Gain, y = fct_reorder(Feature, Gain))) +
  geom_col() +
  labs(title = "Feature Importance", x = "Gain", y = NULL, caption = "LightGBM Model | Visual: Jim Gruman")    

submit_df <-  augment(classification_fit, competition_df) %>%
       select(uid, out = .pred_out, single = .pred_single, double = .pred_double, triple = .pred_triple, home_run = .pred_home_run)

head(submit_df)  %>% 
     knitr::kable()      

submit_df  %>% 
  write_csv("submission.csv")
```    