---
title: "Rainfall"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
    highlight: tango
    code_folding: hide
---
  
# Introduction  {.tabset .tabset-fade .tabset-pills}

The goal of this competition is to predict rainfall for each day of the year.

My notebook serves as a demonstration of some of the possible techniques available to arrive at a solution.  I intend to add to this as I have time available. Your questions and comments are welcome.

Lets dive right in.

The Kaggle kernels have many of the common r packages built in.  

## Load libraries

```{r }
#| label: setup
#| warning: false
#| message: false

if (dir.exists("/kaggle")){
  path <- "/kaggle/input/playground-series-s5e3/"

# options(repos = c(CRAN = "https://packagemanager.posit.co/cran/__linux__/focal/2021-05-17"))
# install.packages("ranger", quiet = TRUE)
    
options(repos = c(CRAN = "https://packagemanager.posit.co/cran/2021-03-22"))
remotes::install_github("tidymodels/plsmod", quiet = TRUE)    
#remotes::install_github("mayer79/outForest", quiet = TRUE)    
#remotes::install_github("tidyverts/fable.binary", quiet = TRUE)
    
cores <- future::availableCores()

} else {
  path <- stringr::str_c(here::here("data"),"/")
  orig_path <- stringr::str_c(here::here("data"),"/")

  cores <- future::availableCores(omit = 2)
}
 
suppressPackageStartupMessages({
library(tidyverse, quietly = TRUE) # metapackage of all tidyverse packages
library(tidymodels) # metapackage see https://www.tidymodels.org/

   
library(plsmod)
library(discrim)

library(DALEXtra) # Tidymodels Explainability         
    
library(brulee)   # interface to torch

})

tidymodels_prefer()
conflicted::conflicts_prefer(brulee::coef)

options(tidymodels.dark = TRUE)

theme_set(ggdark::dark_theme_minimal())

```

## Load Data


```{r }
#| label: load data
#| warning: false
#| message: false

preprocessor <- function(dataframe) {

dataframe <- dataframe %>%
    janitor::clean_names() |>

    mutate(year = id %/% 365 ,
           day = id - year * 365 ) |> 
    
    group_by(day) |> 
    mutate(across(c(dewpoint, maxtemp, pressure, mintemp, temparature, humidity), \(x) (mean(x, na.rm = TRUE) - x)/sd(x, na.rm = TRUE),  .names = "delta_day_{.col}")) |>
    ungroup() |>
    
    mutate(winddirection = factor(winddirection %/% 20)) |> 

    mutate(across(c(where(is.character)), \(x) as.factor(x))) 

return(dataframe)
}

raw_df <- read_csv(str_c(path, "train.csv"),
                   show_col_types = FALSE) |> 
          mutate(rainfall= as.character(rainfall)) |> 
          preprocessor() 

tst_df <- read_csv(str_c(path, "test.csv"),
                   show_col_types = FALSE)  |>  
  preprocessor() 

# because we already know the test set, let's remove the train set factor levels that do not correspond with anything on the test set
for (col in names(raw_df)) {
    if (is.factor(raw_df[[col]]) & col != "rainfall") {
      # Get levels in train and test dataframes
      raw_levels <- levels(raw_df[[col]])
      tst_levels <- levels(tst_df[[col]])
      
      # Identify levels in train not in test
      new_levels <- setdiff(raw_levels, tst_levels)
      
      # Set these levels to NA in train dataframe
      raw_df[[col]] <- factor(raw_df[[col]], levels = c(tst_levels, new_levels))
      raw_df[[col]][raw_df[[col]] %in% new_levels] <- NA_character_
    }
  }

# the synthetic playground competitions seem to perform better when numerics are also included as factors
all_df <-
    bind_rows(raw_df %>% mutate(source = "train"),
              tst_df %>% mutate(source = "test")) 


train_df <- all_df %>% 
  filter(source == "train") %>% 
  select(-source) 

competition_df <- all_df %>% 
  filter(source == "test") %>% 
  select(-source, -rainfall)


```


# EDA {.tabset .tabset-fade .tabset-pills}

## Features

```{r}
#| label: Features
#| warning: false
#| message: false
#| fig.width: 6


features <- train_df %>%
  select(-id, -rainfall, -day, -year, -windspeed) |> 
  names()

train_df <- train_df %>% 
  distinct(pick(all_of(features)), .keep_all = TRUE)

nom_features <- train_df %>%
  select(all_of(features)) %>%
  select(where(is.character), where(is.factor)) %>%
  names() 

logical_features <- train_df %>%
  select(all_of(features)) %>%
  select(where(is.logical)) %>%
  names() 

num_features <- train_df %>%
  select(all_of(features)) %>%
  select(where(is.numeric)) %>%
  names()
```

Nominal features:

`r nom_features`

Numeric features: 

`r num_features`

Logical features: 

`r logical_features`


Size of the combined train and competition datasets:

`r nrow(all_df)`

Size of the split made available to machine learning

`r nrow(train_df)`



## Numeric features

```{r}
#| label: numeric
#| warning: false
#| message: false
#| fig.height: 16
#| fig.width: 6

train_df %>% 
  select(all_of(num_features), rainfall) %>% 
  pivot_longer(-rainfall,
    names_to = "metric",
    values_to = "value"
  ) %>%
  ggplot(aes(value)) +
  geom_histogram(aes(fill = rainfall), position = "identity", bins = 30) +
  geom_freqpoly(aes(linetype = rainfall), bins = 30) +
  geom_rug(aes(color = rainfall)) +
  facet_wrap(vars(metric), scales = "free", ncol = 2) +
  guides(color = "none") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title.position = "plot") +
  labs(color = NULL, 
       title = "Numeric Feature Univariate Distributions",
       caption = "Data: Kaggle.com | Visual: Jim Gruman")

train_df %>% 
  select(all_of(num_features), rainfall, day) %>% 
  pivot_longer(-c(rainfall,day),
    names_to = "metric",
    values_to = "value"
  ) %>%
  ggplot(aes(day, value, color = rainfall)) +
  geom_point(show.legend = FALSE, shape = 20) +
   facet_wrap(vars(metric), scales = "free", ncol = 2) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title.position = "plot") +
  labs(color = NULL, fill = "Rainfall",
       title = "Numeric Feature by Day",
       caption = "Data: Kaggle.com | Visual: Jim Gruman")

```

## Nominal features

Explore the distribution of outcome class by factor level.


```{r}
#| label: nominal
#| warning: false
#| message: false
#| fig.height: 24
#| fig.width: 12


if(length(nom_features) >0){

train_df %>% 
  select(all_of(nom_features), rainfall) %>% 
  mutate(across(nom_features, fct_lump_n,n = 10, other_level = 'other')) %>%

  pivot_longer(-rainfall,
    names_to = "metric",
    values_to = "value"
  ) %>%
    
  summarise(n = n(),
            .by = c(rainfall, metric, value)) %>%

  mutate(value = tidytext::reorder_within(value, n, metric)) %>%
    
  ggplot(aes(x = n, y = value, fill = rainfall)) +
  geom_col() +

  tidytext::scale_y_reordered() +
  scale_x_continuous(n.breaks = 3, guide = guide_axis(n.dodge = 2))  +
  facet_wrap(vars(metric), scales = "free", ncol = 2) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
       legend.position = "bottom") +
  labs(title = "Nominal Feature Counts",
       fill = NULL,
       caption = "Data: Kaggle | Visual: Jim Gruman")
  


train_df %>% 
  select(all_of(nom_features), rainfall) %>% 
  mutate(across(nom_features, fct_lump_n,n = 10, other_level = 'other')) %>%

  pivot_longer(-rainfall,
    names_to = "metric",
    values_to = "value"
  ) %>%
    
  summarise(n = n(),
            .by = c(rainfall, metric, value)) %>%

  mutate(value = tidytext::reorder_within(value, n, metric)) %>%
    
  ggplot(aes(x = n, y = value, fill = rainfall)) +
  geom_col(position = "fill") +

  tidytext::scale_y_reordered() +
  scale_x_continuous(n.breaks = 3, guide = guide_axis(n.dodge = 2))  +
  facet_wrap(vars(metric), scales = "free", ncol = 2) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
       legend.position = "bottom") +
  labs(title = "Proportion of Outcome in Nominal Feature Counts",
       fill = NULL,
       caption = "Data: Kaggle | Visual: Jim Gruman")

}    
```


## Counts of Missingness

The test set is missing one wind direction. We will fill it in with the median of all of the data.
                  
```{r}
#| label: counts of missingness

train_df %>% 
  summarize(across(all_of(features), function(x) sum(is.na(x)))) %>% 
  pivot_longer(everything(),
              names_to = "feature",
              values_to = "Count of Missing") %>% 
                   knitr::kable()

competition_df %>% 
  summarize(across(all_of(features), function(x) sum(is.na(x)))) %>% 
  pivot_longer(everything(),
              names_to = "feature",
              values_to = "Count of Missing") %>% 
                   knitr::kable()

competition_df <- competition_df %>% 
  mutate(across(where(is.numeric), ~replace_na(., median(., na.rm=TRUE))))          

```

## Counts of Distinct
                   
               
```{r}
#| label: counts of distinct
               
train_df %>%
  summarize(across(all_of(features), n_distinct)) %>%
  pivot_longer(everything(), names_to = "feature", values_to = "Count of distinct train") |>
  left_join(
    competition_df %>%
      summarize(across(all_of(features), n_distinct)) %>%
      pivot_longer(everything(), names_to = "feature", values_to = "Count of distinct test")
  ) %>% 
                   knitr::kable()
               
```

## Duplicated

Is this competition transaction already in the training data with a correct rainfall?

```{r}
#| label: duplicates
#| warning: false
#| message: false

bind_rows(train_df %>% mutate(source = "train"),
              competition_df %>% mutate(rainfall= NA_character_, source = "test")) |> 
    group_by_at(features) %>%
    mutate(num_dups = n(),
           dup_id = row_number()) %>% 
    ungroup() %>%
    group_by(source) %>%
    mutate(is_duplicated = dup_id > 1) %>% 
    count(is_duplicated) %>% 
                   knitr::kable()
               

```

## Pairwise Correlations
                   
`ggcorrplot` provides a quick look at numeric features where the correlation may be significant. 

```{r}
#| label: pairwise correlations
#| fig.width: 12
#| fig.height: 12                   
                   
train_df %>% 
  select(all_of(num_features)) %>% 
  ggcorrplot::cor_pmat() %>% 
  ggcorrplot::ggcorrplot(hc.order = TRUE, lab = TRUE,
    type = "lower", insig = "blank") +
  labs(title = "Pairwise Correlations Training Set")
                   
competition_df %>% 
  select(all_of(num_features)) %>% 
  ggcorrplot::cor_pmat() %>% 
  ggcorrplot::ggcorrplot(hc.order = TRUE, lab = TRUE,
    type = "lower", insig = "blank") +
  labs(title = "Pairwise Correlations Competition Set")

```                    
               

## Target

```{r}
#| label: outcome 
#| warning: false
#| message: false
#| fig.width: 6


train_df %>%
  summarize(outcome_sum = n(), .by = rainfall) %>%
  arrange(-outcome_sum) %>%
  mutate(prop = outcome_sum / nrow(train_df)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop) %>%
  ggplot(aes(x = "", y = prop, fill = rainfall)) +
  geom_bar(stat = "identity",
           width = 1,
           show.legend = FALSE) +
  geom_text(
    aes(
      y = ypos,
      label= paste0(rainfall
                     , "\n", round(prop, 2) * 100, "%")
    ),
    color = "white",
    nudge_x = 0,
    size = 3
  ) +
  coord_polar("y", start = 0) +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank()  ) +
  labs(title = "Rainfall", caption = "Data: Kaggle.com | Visual: Jim Gruman")

```                            
           
                
# Machine Learning {.tabset .tabset-fade .tabset-pills}

## Recipe

```{r}
#| label: recipe

base_rec <- recipe(
    
    formula(paste0("rainfall ~ ", 
               str_c(features,  collapse = " + "))),
    data = train_df
  ) %>%     
                   
  step_mutate(cloud_o_sunshine = cloud / (sunshine + 1)) |>   
  step_mutate(humidity_o_dewpoint = humidity / (dewpoint + 1)) |>   

  step_interact(~ humidity:pressure) |>
  step_interact(~ humidity:sunshine) |>
  step_interact(~ cloud:sunshine) |>
  step_interact(~ temparature:humidity) |>                   

  step_impute_mode(all_nominal_predictors()) |>                    
  step_dummy(all_nominal_predictors()) |>     

  step_lag(mintemp, starts_with("delta_day")) |> 

  step_impute_median(all_predictors()) |>                       
                   
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors()) |> 
  themis::step_upsample(rainfall,over_ratio = 1)               


folds <- rsample::sliding_period(
  train_df |>   mutate(id = as.Date("2010-01-01") + day + year*365) ,
  id,
  period = "month",
  lookback = 36L,
  assess_start = 1L,
  assess_stop = 6L,
  complete = TRUE,
  step = 6L,
  skip = 0L
)
                   
folds |>
  tidy() |>
  ggplot(aes(x = Resample, y = factor(Row), fill = Data)) +
  geom_tile() +
  labs(y = "Date") +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
                         
                                     
```



## Workflowset Ensemble

```{r}
#| label: workflowset
#| warning: false
#| message: false
#| fig.width: 12


logistic_reg_glmnet_spec <-
  logistic_reg(penalty = tune(), mixture = tune()) %>%
  set_engine('glmnet')

svm_rbf_kernlab_spec <-
  svm_rbf(cost = tune(), rbf_sigma = tune()) %>%
  set_engine('kernlab') %>%
  set_mode('classification')
                   
pls_mixOmics_spec <-
  pls(predictor_prop = tune(), num_comp = tune()) %>%
  set_engine('mixOmics') %>%
  set_mode('classification')

discrim_linear_mda_spec <-
  discrim_linear(penalty = tune()) %>%
  set_engine('mda')  %>%
  set_mode('classification')

brulee_spec <-
  mlp(penalty = tune(), hidden_units = 5L, epochs = 1000L) %>%
  set_engine('brulee', stop_iter = 100L, rate_schedule = "cyclic")   %>%
  set_mode('classification')     
          

dep_models <- 
   workflow_set(
      preproc = list(base = base_rec),
      models = list(linear = logistic_reg_glmnet_spec,
                    svm = svm_rbf_kernlab_spec,
                    pls = pls_mixOmics_spec,
                    discrim = discrim_linear_mda_spec,
                    brulee = brulee_spec),
      cross = TRUE
   ) %>% 
  option_add_parameters() |> 
  option_add(
    control = stacks::control_stack_grid(),
    metrics = metric_set(mn_log_loss)
  )

glm_params <- dep_models |> 
    extract_workflow("base_linear") |> 
  parameters() |> 
  update(
    mixture = mixture(),
    penalty = dials::penalty(range = c(-1,0)))
                  
svm_params <- dep_models |> 
    extract_workflow("base_svm") |> 
  parameters() |> 
  update(
    cost = cost(),
    rbf_sigma = dials::rbf_sigma(range = c(-4,-8)))
       
pls_params <- dep_models |>
       extract_workflow("base_pls") |>       
  parameters() |> 
  update(
    num_comp = num_comp(range = c(2L,3L)),
    predictor_prop = predictor_prop(range = c(0,0.25)))

brulee_params <- dep_models |>
       extract_workflow("base_brulee") |>       
  parameters() |> 
  update(
    penalty = penalty(range = c(-2,0)))       
                   
dep_models <- dep_models |> 
  option_add(
    param_info = svm_params,
    id = "base_svm"
  ) |> 
  option_add(
    param_info = glm_params,
    id = "base_linear"
  ) |> 
  option_add(
    param_info = pls_params,
    id = "base_pls"
  ) |> 
  option_add(
    param_info = brulee_params,
    id = "base_brulee"
  ) |> 
  option_add(
    param_info = brulee_params,
    id = "base_brulee"
  ) |> 
   workflow_map("tune_grid", resamples = folds, grid = 25, 
                metrics = metric_set(roc_auc), verbose = TRUE)

rank_results(dep_models, rank_metric = "roc_auc", select_best = TRUE)                    

autoplot(dep_models) +
  geom_text(aes(y = mean -0.03, label= wflow_id), angle = 90, hjust = 1)+
  theme(legend.position = "none")

```

## Hyperparameters and Feature Importance

```{r }                   
#| label: linear Details
#| warning: false
#| message: false                   
                   

dep_models %>%
  dplyr::filter(grepl("base_linear", wflow_id)) %>%
  dplyr::mutate(metrics = purrr::map(result, tune::collect_metrics)) %>%
  dplyr::select(wflow_id, metrics) %>%
  tidyr::unnest(cols = metrics) |> 
  arrange(desc(mean))

dep_models |> 
  workflowsets::extract_workflow_set_result("base_linear") |> 
  autoplot() +
  labs(title = "GLMNET Hyperparameter Search")

best_linear_params <- dep_models |> 
  workflowsets::extract_workflow_set_result("base_linear") |> 
  select_best(metric = "roc_auc")  

dep_models |> 
  workflowsets::extract_workflow("base_linear") |> 
  finalize_workflow(best_linear_params) |> 
  fit(train_df) |> 
  vip::vip(num_features = 30L)                   
                                
                   
                   
```

                   
```{r }                   
#| label: svm Details
#| warning: false
#| message: false 
                   
dep_models %>%
  dplyr::filter(grepl("base_svm", wflow_id)) %>%
  dplyr::mutate(metrics = purrr::map(result, tune::collect_metrics)) %>%
  dplyr::select(wflow_id, metrics) %>%
  tidyr::unnest(cols = metrics) |> 
  arrange(desc(mean))

dep_models |> 
  workflowsets::extract_workflow_set_result("base_svm") |> 
  autoplot() +
  labs(title = "SVM Hyperparameter Search")                   

       
                    

```


```{r }                   
#| label: pls Details
#| warning: false
#| message: false                    
dep_models %>%
  dplyr::filter(grepl("base_pls", wflow_id)) %>%
  dplyr::mutate(metrics = purrr::map(result, tune::collect_metrics)) %>%
  dplyr::select(wflow_id, metrics) %>%
  tidyr::unnest(cols = metrics) |> 
  arrange(desc(mean))

dep_models |> 
  workflowsets::extract_workflow_set_result("base_pls") |> 
  autoplot() +
  labs(title = "PLS Hyperparameter Search")                        
                      
                   
```

```{r }                   
#| label: discrim Details
#| warning: false
#| message: false                    
dep_models %>%
  dplyr::filter(grepl("base_discrim", wflow_id)) %>%
  dplyr::mutate(metrics = purrr::map(result, tune::collect_metrics)) %>%
  dplyr::select(wflow_id, metrics) %>%
  tidyr::unnest(cols = metrics) |> 
  arrange(desc(mean))

dep_models |> 
  workflowsets::extract_workflow_set_result("base_discrim") |> 
  autoplot() +
  labs(title = "Discrim Hyperparameter Search")   
                   
```          

```{r }                   
#| label: brulee Details
#| warning: false
#| message: false                    
dep_models %>%
  dplyr::filter(grepl("base_brulee", wflow_id)) %>%
  dplyr::mutate(metrics = purrr::map(result, tune::collect_metrics)) %>%
  dplyr::select(wflow_id, metrics) %>%
  tidyr::unnest(cols = metrics) |> 
  arrange(desc(mean))

dep_models |> 
  workflowsets::extract_workflow_set_result("base_brulee") |> 
  autoplot() +
  labs(title = "Brulee Hyperparameter Search")   
                   
```           

## Select Best                

```{r }                   
#| label: Best Model fit
#| warning: false
#| message: false 

best_model_id <- dep_models |> 
  rank_results(
    rank_metric = "roc_auc",
    select_best = TRUE
  ) |> 
  head(1) |> 
  pull(wflow_id)

best_model_id

best_model <- extract_workflow(dep_models, id = best_model_id)

best_workflow <- 
  dep_models |> 
  extract_workflow_set_result(id = best_model_id)

best_workflow                   

select_best(best_workflow, metric =   "roc_auc")

final_wf <- finalize_workflow(best_model, select_best(best_workflow,  metric =   "roc_auc"))
                 
classification_fit <- fit(final_wf, data = train_df)

```

# Performance {.tabset .tabset-fade .tabset-pills}

## DALEX

```{r}
#| label: explainer
#| warning: false
#| message: false

explainer <- 
  explain_tidymodels(
    classification_fit, 
    data = train_df %>% dplyr::select(all_of(features)), 
    y = as.numeric(train_df$rainfall),
    label = "Ensemble",
    verbose = FALSE
  )  %>% 
  model_parts()

ggplot_imp <- function(...) {
  obj <- list(...)
  metric_name <- attr(obj[[1]], "loss_name")
  metric_lab <- paste(metric_name, 
                      "after permutations\n(higher indicates more important)")
  
  full_vip <- bind_rows(obj) %>%
    filter(variable != "_baseline_")
  
  perm_vals <- full_vip %>% 
    filter(variable == "_full_model_") %>% 
    group_by(label) %>% 
    summarise(dropout_loss = mean(dropout_loss))
  
  p <- full_vip %>%
    filter(variable != "_full_model_") %>% 
    mutate(variable = fct_reorder(variable, dropout_loss)) %>%
    ggplot(aes(dropout_loss, variable)) 
  if(length(obj) > 1) {
    p <- p + 
      facet_wrap(vars(label)) +
      geom_vline(data = perm_vals, aes(xintercept = dropout_loss, color = label),
                 linewidth = 1.4, lty = 2, alpha = 0.7) +
      geom_boxplot(aes(color = label, fill = label), alpha = 0.2)
  } else {
    p <- p + 
      geom_vline(data = perm_vals, aes(xintercept = dropout_loss),
                 linewidth = 1.4, lty = 2, alpha = 0.7) +
      geom_boxplot(fill = "#91CBD765", alpha = 0.4)
    
  }
  p +
    theme(legend.position = "none") +
    labs(x = metric_lab, 
         y = NULL,  fill = NULL,  color = NULL)
}
                   
ggplot_imp(explainer)     
                   

```                   

## Submission
```{r }                   
#| label: submission
#| warning: false
#| message: false
                   
augment(classification_fit, train_df) %>% 
  conf_mat(rainfall, .pred_class) %>%
  yardstick:::autoplot.conf_mat(type = "heatmap")
                   
augment(classification_fit, train_df, type = "prob") %>%
    yardstick::roc_curve(rainfall, .pred_0) |> 
    ggplot(aes(1-specificity, sensitivity)) +
    geom_line() +
    geom_abline(slope = 1) +
    coord_equal() 
                   
submit_df <-  augment(classification_fit , competition_df, type= "prob") %>%
       transmute(id = round(id), rainfall = .pred_1)

head(submit_df)  %>% 
     knitr::kable()      

submit_df  %>% 
  write_csv("submission.csv")
```  